Predicting Survival Probability of NASA Aircraft Engines

### 2025-04-27 index.qmd
updates

ggsurvplot(example_curve,
           data = lung,
           title = "Example Survival Curve",
           xlab = "Time",
           ylab = "Survival Probability",
           legend = "none")
```

@@

plot_model(forest,
           dot.size = 1,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")

@@

autoplot(km_fit) + theme_bw()
```

@@

```{r, echo = FALSE}
plot_model(cox1.1,
           dot.size = 2,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")
```

@@

```{r, echo = FALSE}
par(mfrow = c(4,4),
      mar = c(2,2,4,1))

plot(data_grouped$os1, martingale, main = "Martingale Residuals: os1")
abline(h = 0, col = "red")

plot(data_grouped$os2, martingale, main = "Martingale Residuals: os2")
abline(h = 0, col = "red")

plot(data_grouped$os3, martingale, main = "Martingale Residuals: os3")
abline(h = 0, col = "red")

plot(data_grouped$sm1, martingale, main = "Martingale Residuals: sm1")
abline(h = 0, col = "red")

plot(data_grouped$sm2, martingale, main = "Martingale Residuals: sm2")
abline(h = 0, col = "red")

plot(data_grouped$sm3, martingale, main = "Martingale Residuals: sm3")
abline(h = 0, col = "red")

plot(data_grouped$sm4, martingale, main = "Martingale Residuals: sm4")
abline(h = 0, col = "red")

plot(data_grouped$sm5, martingale, main = "Martingale Residuals: sm5")
abline(h = 0, col = "red")

plot(data_grouped$sm6, martingale, main = "Martingale Residuals: sm6")
abline(h = 0, col = "red")

plot(data_grouped$sm7, martingale, main = "Martingale Residuals: sm7")
abline(h = 0, col = "red")

plot(data_grouped$sm8, martingale, main = "Martingale Residuals: sm8")
abline(h = 0, col = "red")

plot(data_grouped$sm9, martingale, main = "Martingale Residuals: sm9")
abline(h = 0, col = "red")

plot(data_grouped$sm10, martingale, main = "Martingale Residuals: sm10")
abline(h = 0, col = "red")

plot(data_grouped$sm11, martingale, main = "Martingale Residuals: sm11")
abline(h = 0, col = "red")

plot(data_grouped$sm12, martingale, main = "Martingale Residuals: sm12")
abline(h = 0, col = "red")

plot(data_grouped$sm13, martingale, main = "Martingale Residuals: sm13")
abline(h = 0, col = "red")

plot(data_grouped$sm14, martingale, main = "Martingale Residuals: sm14")
abline(h = 0, col = "red")

plot(data_grouped$sm15, martingale, main = "Martingale Residuals: sm15")
abline(h = 0, col = "red")

plot(data_grouped$sm17, martingale, main = "Martingale Residuals: sm17")
abline(h = 0, col = "red")

plot(data_grouped$sm18, martingale, main = "Martingale Residuals: sm18")
abline(h = 0, col = "red")

plot(data_grouped$sm20, martingale, main = "Martingale Residuals: sm20")
abline(h = 0, col = "red")

plot(data_grouped$sm21, martingale, main = "Martingale Residuals: sm21")
abline(h = 0, col = "red")
```

@@

create_ph_plots(cox1.1, 3, 5)
```

@@

ggsurvplot(fit = sf,
           data = data_grouped_reduced, risk.table = TRUE, fun = "cumhaz")$plot +
  ggtitle('Cumulative Hazard Over Time')

@@

ggsurvplot(fit = sf, data = data_grouped_reduced)$plot +
  ggtitle('Survival Probability Over Time')

@@

The lower end of the confidence interval for survival probability is
shown to hold stable at 100% until `t = 128`, with a gradual decrease
afterwards. Half of the engines are expected to survive until `t = 245`,
and 90% of the engines are expected to fail at `t = 347`. These results
can be viewed in both the above survival graph and the cumulative hazard
plot which shows a low hazard until `t = 128` with a gradual increase.

@@

>>>>

ggsurvplot(example_curve,
           data = lung,
           title = "Example Survival Curve",
           xlab = "Time",
           ylab = "Survival Probability (%)",
           legend = "none")
```

@@

plot_model(forest,
           dot.size = 1,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  xlab(label = "Covariates")+
  ylab(label = "Estimated Effect on Hazard Ratio")

@@

autoplot(km_fit) + theme_bw() + labs(y = "Survival Probability (%)", x = 'Time (Iterations)', title = "Kaplan-Meier Survival Curve")
```

@@

```{r, echo = FALSE}
plot_model(cox1.1,
           dot.size = 2,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratio")+
  xlab(label = "Covariates")+
  ylab(label = "Estimated Effect on Hazard Ratio")
```

@@

```{r, echo = FALSE}
par(mfrow = c(4,6),
      mar = c(2,2,4,1))

plot(data_grouped$os1, martingale, main = "Martingale: os1")
abline(h = 0, col = "red")

plot(data_grouped$os2, martingale, main = "Martingale: os2")
abline(h = 0, col = "red")

plot(data_grouped$os3, martingale, main = "Martingale: os3")
abline(h = 0, col = "red")

plot(data_grouped$sm1, martingale, main = "Martingale: sm1")
abline(h = 0, col = "red")

plot(data_grouped$sm2, martingale, main = "Martingale: sm2")
abline(h = 0, col = "red")

plot(data_grouped$sm3, martingale, main = "Martingale: sm3")
abline(h = 0, col = "red")

plot(data_grouped$sm4, martingale, main = "Martingale: sm4")
abline(h = 0, col = "red")

plot(data_grouped$sm5, martingale, main = "Martingale: sm5")
abline(h = 0, col = "red")

plot(data_grouped$sm6, martingale, main = "Martingale: sm6")
abline(h = 0, col = "red")

plot(data_grouped$sm7, martingale, main = "Martingale: sm7")
abline(h = 0, col = "red")

plot(data_grouped$sm8, martingale, main = "Martingale: sm8")
abline(h = 0, col = "red")

plot(data_grouped$sm9, martingale, main = "Martingale: sm9")
abline(h = 0, col = "red")

plot(data_grouped$sm10, martingale, main = "Martingale: sm10")
abline(h = 0, col = "red")

plot(data_grouped$sm11, martingale, main = "Martingale: sm11")
abline(h = 0, col = "red")

plot(data_grouped$sm12, martingale, main = "Martingale: sm12")
abline(h = 0, col = "red")

plot(data_grouped$sm13, martingale, main = "Martingale: sm13")
abline(h = 0, col = "red")

plot(data_grouped$sm14, martingale, main = "Martingale: sm14")
abline(h = 0, col = "red")

plot(data_grouped$sm15, martingale, main = "Martingale: sm15")
abline(h = 0, col = "red")

plot(data_grouped$sm17, martingale, main = "Martingale: sm17")
abline(h = 0, col = "red")

plot(data_grouped$sm18, martingale, main = "Martingale: sm18")
abline(h = 0, col = "red")

plot(data_grouped$sm20, martingale, main = "Martingale: sm20")
abline(h = 0, col = "red")

plot(data_grouped$sm21, martingale, main = "Martingale: sm21")
abline(h = 0, col = "red")
```

@@

create_ph_plots(cox1.1, 4, 6)
```

@@

ggsurvplot(fit = sf,
           data = data_grouped_reduced, risk.table = TRUE, fun = "cumhaz")$plot +
  ggtitle('Cumulative Hazard Over Time')+
  xlab("Time (Iterations)")

@@

ggsurvplot(fit = sf, data = data_grouped_reduced)$plot +
  ggtitle('Survival Probability Over Time')+
  xlab('Timee (Iterations)')+
  ylab('Survival Probability (%)')

@@

The lower end of the confidence interval for survival probability is
shown to hold stable at 100% until `t = 128`, with a gradual decrease
afterwards. Half of the engines are expected to survive until `t = 245`,
and 90% of the engines are expected to fail at `t = 347`. These results
can be viewed in both the above survival graph and the cumulative hazard
plot which shows a low hazard until `t = 128` with a gradual increase. The model's final concordance index of .696 indicates that the calculated confidence interval is fairly accurate, but it is still subject to a substantial margin for error.

@@

While the concordance index of .696 achieved in this analysis demonstrates the model is fairly accurate, there is still room for improvement. Survival models resulting in a concordance index of .9 and higher are highly accurate and would provide a tighter, more accurate confidence interval for the longevity of engines. After attempting transformations of the data and creating models many different combinations of covariates, the final c-index that was calculated is the highest value achieved where the covariates were able to meet the proportional hazards model.

As discovered during the literature review, a weakness of the cox proportional hazards model is the necessity to meet the proportional hazards assumption. Due to the difficulty meeting the assumption during this analysis, to achieve a more accurate model it would be beneficial to explore other survival models for this dataset.
